[Re] Marriage and happiness: Providing evidence against a relationship between inequality and happiness in Oishi, Kesebir, and Diener (2011)

### 2020-02-21 analysis/paper.Rmd
corrected fig3 issues and both replicated and reproduced it

# ==== Clean data
# Subset the GSS data to just the relevant fields
gss <- gss_all %>%
  select(year, happy, race, marital) %>%
  mutate(year = as.integer(year), # fix data class
         happy = as.integer(happy),
         happy = recode(happy, `1`= 3, `2` = 2, `3` = 1), # reverse the values such that higher is more happy
         race = as.integer(race),
         marital = as.integer(marital)) %>%
  na.omit() # remove NAs (4780/64,814 cases where not all data was collected)

# join the gini and happiness/race/marital data
happy <- gini %>%
  left_join(gss, by = c("year" = "year")) %>%
  na.omit()

# create summary table by year
happy_race_summary <- happy %>%
  group_by(year, race) %>%
  summarize(happy_avg = mean(happy))

@@

```{r plots}
g <- ggplot(happy_race_summary, aes(x = year, y = happy_avg, color = factor(race)))
g + geom_point() +
  geom_line(stat = "smooth",
            method = "lm", 
            se = FALSE, 
            alpha = 0.5) +
  theme_bw() +
  labs(title = "Figure 3",
       x = "Year",
       y = "Mean Happiness",
       caption = "Figure 3") +
  scale_color_discrete(name="Race",
                      breaks=c(1, 2, 3),
                      labels=c("White", "Black", "Other")) +
  scale_y_continuous(limits = c(1.9,2.5))

>>>>

# ==== Clean data
# Subset the GSS data to just the relevant fields for the relevant time frame, join, and recode the same way the 
  # original study does
study_data <- gss_all  %>%
  select(year, happy, race, marital) %>%
  filter(year <= 2012) %>%
  mutate(year = as.integer(year),
         happy = dplyr::recode(happy, `1`= 3L, `2` = 2L, `3` = 1L)) %>% #recode to increasing happiness from 1 to 3
  left_join(gini, by = c("year" = "year"))

# Relabel the labels attribute to reflect the inverse of the happiness intensity (not actually used, but prolly good)
attributes(study_data$happy)$labels <- c("NOT TOO HAPPY" = 1, "PRETTY HAPPY" = 2, "VERY HAPPY" = 3)

# Create summary table of mean happiness by year by race
race_summary <- study_data %>%
  group_by(year, race) %>%
  dplyr::summarize(happy_avg = mean(na.omit(happy))) # we can remove the 'dplyr' once we only have the tidyverse loaded

@@

```{r plots}

# Fig 3 reproduction
g <- ggplot(race_summary, aes(x = year, y = happy_avg, color = factor(race)))
g + geom_point(size = .6) +
  geom_line(stat = "smooth",
            method = "lm", 
            se = FALSE, 
            alpha = 0.5) +
  theme_bw() +
  labs(title = "Average Happiness by Race Trend",
       x = "Year",
       y = "Mean Happiness",
       caption = "Figure 3") +
  scale_color_manual(values=c("blue", "forestgreen", "dodgerblue"),
                     name="Race",
                     breaks=c(1, 2, 3),
                     labels=c("White", "Black", "Other")) +
  coord_cartesian(ylim = c(1.9,2.5)) # setting this to match the study excludes the lm via the above methodology
